Weighted Model Counting/Integration from the Perspective of Boolean Algebras

### 2020-03-20 paper.tex
First version of the introductory example

\begin{figure}
  \[
    \begin{tikzcd}
      & & & & & \colorbox{color3}{$1$} \ar[dlll,dash,gray] \ar[dl,dash,gray]
      \ar[dr,dash,gray] \ar[drrr,dash,gray] & & & \\
      & & \colorbox{color3}{$p \lor q$} & & \colorbox{color3}{$q \to p$} & &
      \colorbox{color2}{$p \to q$} & & \colorbox{color4}{$\neg p \lor \neg q$}
      \\
      \colorbox{color3}{$p$} \ar[urr,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$q$} \ar[u,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$p \leftrightarrow q$} \ar[u,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$p + q$} \ar[ullll,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$\neg q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] & & \colorbox{color1}{$\neg p$} \ar[ullll,dash,gray]
      \ar[ull,dash,gray] \\
      & & \fcolorbox{black}{color2}{$p \land q$} \ar[ull,dash,gray]
      \ar[u,dash,gray] \ar[urr,dash,gray] & & \fcolorbox{black}{color4}{$p \land
        \neg q$} \ar[ullll,dash,gray] \ar[urr,dash,gray] \ar[urrrr,dash,gray] &
      & \fcolorbox{black}{color1}{$\neg p \land q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] \ar[urrrr,dash,gray] & & \fcolorbox{black}{color1}{$\neg
        p \land \neg q$} \ar[ullll,dash,gray] \ar[u,dash,gray]
      \ar[urr,dash,gray] \\
      & & & & & \colorbox{color1}{$0$} \ar[ulll,dash,gray] \ar[ul,dash,gray]
      \ar[ur,dash,gray] \ar[urrr,dash,gray] & & &
    \end{tikzcd}
  \]
  \[
    \begin{tikzcd}
      & \colorbox{color3}{$\left[1\right]$} \ar[dl,dash,gray] \ar[dr,dash,gray]
      & \\
      \fcolorbox{black}{color2}{$\left[q\right]$} & &
      \fcolorbox{black}{color4}{$\left[\neg q\right]$} \\
      & \colorbox{color1}{$\left[0\right]$} \ar[ul,dash,gray] \ar[ur,dash,gray]
      &
    \end{tikzcd}
  \]
  \caption{Two BAs from \cref{example:construction}: $B(\mathcal{L})$ at the top
    and $B(\Delta)$ at the bottom. An edge between elements $a$ and $b$ (with
    $a$ positioned lower than $b$) means that $a \le b$. Each element of
    $B(\Delta)$ is an equivalence class of elements of $B(\mathcal{L})$, and the
    colours show which elements of $B(\mathcal{L})$ belong to which class. In
    both algebras, atoms have borders around them.}
  \label{fig:example}
\end{figure}

@@

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$---this corresponds to $B(\mathcal{L})$ in
  \cref{fig:example}---and then take a quotient with respect to either the
  filter generated by $p$ or the ideal\footnote{More details on these concepts
    can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$. In any case, the resulting BA is pictured at the bottom of
  \cref{fig:example}. To summarise, 
\end{example}
% TODO: could talk about how each element can be seen as a set of models.
% TODO: the number of elements in the algebra.

% TODO: Feedback: Probably a good idea to explain how elements of the Boolean
% algebra map to models in WMC formulations. Probably do an example:
% 0.3 p; 0.2 q => M = <p, qbar> has weight 0.3 * (1-0.2)
% then it corresponds to an atom in an algebra defined over the language defined
% over propositions p and q.
% TODO: connect the two definitions of WMC

>>>>

\begin{figure}
  \[
    \begin{tikzcd}
      & & & & & \colorbox{color3}{$1$} \ar[dlll,dash,gray] \ar[dl,dash,gray]
      \ar[dr,dash,gray] \ar[drrr,dash,gray] & & & \\
      & & \colorbox{color3}{$p \lor q$} & & \colorbox{color3}{$q \to p$} & &
      \colorbox{color2}{$p \to q$} & & \colorbox{color4}{$\neg p \lor \neg q$}
      \\
      \colorbox{color3}{$p$} \ar[urr,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$q$} \ar[u,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$p \leftrightarrow q$} \ar[u,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$p + q$} \ar[ullll,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$\neg q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] & & \colorbox{color1}{$\neg p$} \ar[ullll,dash,gray]
      \ar[ull,dash,gray] \\
      & & \fcolorbox{black}{color2}{$p \land q$} \ar[ull,dash,gray]
      \ar[u,dash,gray] \ar[urr,dash,gray] & & \fcolorbox{black}{color4}{$p \land
        \neg q$} \ar[ullll,dash,gray] \ar[urr,dash,gray] \ar[urrrr,dash,gray] &
      & \fcolorbox{black}{color1}{$\neg p \land q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] \ar[urrrr,dash,gray] & & \fcolorbox{black}{color1}{$\neg
        p \land \neg q$} \ar[ullll,dash,gray] \ar[u,dash,gray]
      \ar[urr,dash,gray] \\
      & & & & & \colorbox{color1}{$0$} \ar[ulll,dash,gray] \ar[ul,dash,gray]
      \ar[ur,dash,gray] \ar[urrr,dash,gray] & & &
    \end{tikzcd}
  \]
  \[
    \begin{tikzcd}
      & \colorbox{color3}{$\left[1\right]$} \ar[dl,dash,gray] \ar[dr,dash,gray]
      & \\
      \fcolorbox{black}{color2}{$\left[q\right]$} & &
      \fcolorbox{black}{color4}{$\left[\neg q\right]$} \\
      & \colorbox{color1}{$\left[0\right]$} \ar[ul,dash,gray] \ar[ur,dash,gray]
      &
    \end{tikzcd}
  \]
  \caption{Two BAs from \cref{example:construction}: $B(\mathcal{L})$ at the top
    and $B(\Delta)$ at the bottom. An edge between elements $a$ and $b$ (with
    $a$ positioned lower than $b$) means that $a < b$. Each element of
    $B(\Delta)$ is an equivalence class of elements of $B(\mathcal{L})$, and the
    colours show which elements of $B(\mathcal{L})$ belong to which class. In
    both algebras, atoms have borders around them.}
  \label{fig:example}
\end{figure}

@@

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$---this corresponds to $B(\mathcal{L})$ in
  \cref{fig:example}---and then take a quotient with respect to either the
  filter generated by $p$ or the ideal\footnote{More details on these concepts
    can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$. In any case, the resulting BA is pictured at the bottom of
  \cref{fig:example}.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate} % TODO: why is step 1 always possible?
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
% TODO: how to compute the number of elements in the algebra.
% TODO: clarify what B(L) means. And whether B(Delta) is even necessary.
% TODO: reference for the set/subset thing.
